% FILE: acknowledgement.tex  Version 2.1
% AUTHOR:
% Universität Duisburg-Essen, Standort Duisburg
% AG Prof. Dr. Günter Törner
% Verena Gondek, Andy Braune, Henning Kerstan
% Fachbereich Mathematik
% Lotharstr. 65., 47057 Duisburg
% entstanden im Rahmen des DFG-Projektes DissOnlineTutor
% in Zusammenarbeit mit der
% Humboldt-Universitaet zu Berlin
% AG Elektronisches Publizieren
% Joanna Rycko
% und der
% DNB - Deutsche Nationalbibliothek

\section*{Acknowledgment}
\addcontentsline{toc}{section}{Acknowledgment}

At the beginning of this thesis, I would like to say a really big thank you to
all people who supported me throughout these difficult but immensely enjoyable
years of work on my dissertation; to people who always were there, when I
needed them; to people who taught, advised, and helped me with my research and
who, I believe, made me stronger, perhaps more clever, and probably also more
confident in my knowledge.  These people are my colleagues, my friends, my
teachers, but, first of all, my family and my advisor.

I am heavily indebted to all my supporters and I am absolutely convinced that
this work had never come into existence without them.  I am also fairly sure
that it would certainly take me another dissertation to express the whole debt
of gratitude that I owe each of my helpers, but I am afraid they would not
like to wait for me that long to finish it.  Therefore, as hard as it is, I
have to keep myself short in this acknowledgment and should try to describe
all the infinite gratefulness that I feel with just few but nevertheless
utterly sincere words.

First and foremost, I, of course, would like to thank my advisor, Professor
Manfred Stede, who was much more a ``Doktorvater'' for me than just a
supervisor.  I thank Manfred for his courage in employing me, for the numerous
hours that he invested in our discussions, and for his valuable hints,
suggestions, and criticism.  His experience and wisdom laid the foundation for
the success of this work.  And his patience allowed me to bring this thesis to
the end.  I am very thankful to Manfred that he introduced me to the amazing
world of science, a world which does not require a visa but which does require
a sober mind and a considerable load of knowledge.  I sincerely hope that he
will open the doors to this unbelievably beautiful world for many other
Ph.D. students that he will certainly have in the future.

For sure, having a good ``Doktorvater'' is an obligatory precondition for any
completed dissertation, but another prerequisite which plays an equally
important role is still one's own family at home.  And I have to admit that I
was extremely lucky with both my families: the scientific and the traditional
one.  My wife and my parents have always given me incredible support.  But,
unfortunately, they also were those who, I think, suffered most from my
absence and lack of my spare time.  So, taking this rare opportunity to speak
to them from these pages, I would like to simultaneously apologize for my
possibly neglected family duties and to promise that I will make up for them
twofold.

Of course, my time at the University of Potsdam had never been that enjoyable
if I did not have such great colleagues as Tatjana Scheffler, Stephan Oepen,
Arne Neumann, Yulia Grishina, Andreas Peldszus, and Jonathan Sonntag.  I thank
Tatjana for her help with our research project ``Discourse Analysis of Social
Media'' and for her valuable advices and corrections of my papers.  I thank
Arne for our talks, cultural excursions, and his nice attitude to people.  And
I am, of course, much indebted to Yulia who was the most reliable annotator in
our numerous conducted annotation experiments.

I am also grateful to our colleagues from the chair of Theoretical
Computational Linguistics -- Alexander Koller and Thomas Hanneforth.  The
unsurpassable level of Alexander's competence and the consummate quality of
his work set an example for me that I eagerly tried to follow.  And even if my
attempt is still far away from being completed, it would have been even less
successful if I did not take the unique opportunity to visit Thomas
Henneforth's courses in automata theory and C++.  And I am especially thankful
to Thomas for his time and his priceless research hints that he gave me in our
consultations.

I also want to express my gratitude to Konstantina Garoufi, Nikos
Engonopoulos, Mart\'{i}n Villalba, and \v{Z}eljko Agi\'{c}.  It has always
been a pleasure for me to listen to them in our talks and to enjoy their
company during our lunches.  I wish them great success in their academic and
industrial careers and I am also already looking forward to seeing them at
conferences, workshops, and meetings.

Last but not least, I would like to thank our technical assistances: Annett
E\ss{}linger , Ralf Ditsch, and Siegfried Wrobel for their help in solving
everyday issues and problems.  Annett has always been indispensable in
handling all my questions concering the university officials.  And thanks to
Ralf and Siegfried, we had an excellent network infrastructure which helped me
organize my work more efficiently.

There are, of course, many more people whom I would like to mention and who
made a significant contribution to the progress of this thesis, but, as I
said, I have to keep myself short.  So, if I failed to mention somebody on
these pages, I can assure that I did that not due to the lack of my gratitude
but due to the lack of writing space and time.  And I would like to address my
final big thank you in this acknowledgment to all those who remained unnamed
in this section but who feel that they have made this work possible.

\begin{flushright}
\textit{U.S.}
\end{flushright}
